Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations

### 2020-05-06 paper/tidy-missing-data-paper.Rmd
cleaning up, up to section 3.2

\pkg{VIM} [@VIM] implements well-used imputation methods K nearest neighbors, regression, hot-deck, and iterative robust model-based imputation. These diverse approaches allows for imputing with semi-continuous, continuous, count, and categorical data. VIM identifies imputed cases by adding an indicator variable with a suffix `_imp`. So `Var1` has a sibling column, `Var1_imp`, with values TRUE or FALSE indicating imputation. \pkg{VIM} also has a variety of visualization methods, discussed in \@ref(exploration). \pkg{simputation} provides an interface to imputation methods from VIM, in addition providing hotdeck imputation, and the EM algorithm [@simputation; @Dempster1977]. \pkg{simputation} provides a consistent formula interface to all imputation, and always returns a dataframe with the updated imputations. \pkg{Hmisc} [@Hmisc], provide predictive mean matching, \pkg{imputeTS} [@imputeTS], provides time series imputation methods, and \pkg{missMDA} [@missMDA], imputes data using principal components analysis.

@@

\pkg{Amelia} [@amelia] assumes data are multivariate normal, and samples from the posterior using the computationally efficient (and parallelizable) Expectation-Maximization Bootstrap (EMB) algorithm [@Honaker2010] and allows for incorporation of information on the values in a prior. \pkg{norm} [@norm], provides multiple imputation using EM for multivariate normal data, drawing from methods in the NORM software [@schafer-norm]. \pkg{norm} does not provide a framework for tracking missing values, instead providing tools for making inference from multiple imputation.

\pkg{mi} [@mi] also uses Bayesian models for imputation, providing better handling of semi-continuous values, and data with structural or perfect correlation. A collection of analysis models are also provided in \pkg{mi}, which work with the specially multiply imputed data, including linear models, generalized linear models, and their corresponding Bayesian components. This approach promotes fluid workflow, with a similar penalty to tidying up model output, which is still messy.

@@

Each imputation method provides practical methods for different usecases, but most have different output structures, and do not have consistent interfaces in their implementation. This makes then inherently messy and challenging to integrate into an analysis pipeline. For example, combining different imputation methods from different pieces of software is not currently straightforward. \pkg{simputation} is different from other imputation packages, and resolves some of these complications with a simple approach of a unified syntax for all imputation, and always returning a dataframe containing imputed values. This reduces the friction of working with other tools, but comes at the cost of identifying imputed values. An ideal approach would use consistent, simple data structures that work with other analysis tools, and help track missing values. This would make imputation outputs tidy, streamlining subsequent analysis.

>>>>

\pkg{VIM} [@VIM] implements well-used imputation methods K nearest neighbors, regression, hot-deck, and iterative robust model-based imputation. These diverse approaches allows for imputing with semi-continuous, continuous, count, and categorical data. VIM identifies imputed cases by adding an indicator variable with a suffix `_imp`. So `Var1` has a sibling column, `Var1_imp`, with values TRUE or FALSE indicating imputation. \pkg{VIM} also has a variety of visualization methods, discussed in \@ref(exploration). \pkg{simputation} provides an interface to imputation methods from VIM, in addition providing hotdeck imputation, and the EM algorithm [@simputation; @Dempster1977]. \pkg{simputation} provides a consistent formula interface for all imputation methods, and always returns a dataframe with the updated imputations. \pkg{Hmisc} [@Hmisc], provides predictive mean matching, \pkg{imputeTS} [@imputeTS], provides time series imputation methods, and \pkg{missMDA} [@missMDA], imputes data using principal components analysis.

@@

\pkg{Amelia} [@amelia] assumes data are multivariate normal, and samples from the posterior, and allows for incorporation of information on the values in a prior. It uses the computationally efficient (and parallelizable) Expectation-Maximization Bootstrap (EMB) algorithm [@Honaker2010]. \pkg{norm} [@norm], provides multiple imputation using EM for multivariate normal data, drawing from methods in the NORM software [@schafer-norm]. \pkg{norm} does not provide a framework for tracking missing values, instead providing tools for making inference from multiple imputation.

\pkg{mi} [@mi] also uses Bayesian models for imputation, providing better handling of semi-continuous values, and data with structural or perfect correlation. A collection of analysis models are also provided in \pkg{mi}, to work with data it has imputed. These include linear models, generalized linear models, and their corresponding Bayesian components. This approach promotes fluid workflow, with a similar penalty to tidying up model output, which is still messy.

@@

Each imputation method provides practical methods for different usecases, but most have different output structures, and do not have consistent interfaces in their implementation. This makes them inherently messy and challenging to integrate into an analysis pipeline. For example, combining different imputation methods from different pieces of software is not currently straightforward. \pkg{simputation} is different from other imputation packages, and resolves some of these complications with a simple approach of a unified syntax for all imputation, and always returning a dataframe containing imputed values. This reduces the friction of working with other tools, but comes at the cost of identifying imputed values. An ideal approach would use consistent, simple data structures that work with other analysis tools, and help track missing values. This would make imputation outputs tidy, streamlining subsequent analysis.
